Network Deliberation: The role of network structure in large-scale, internet-enabled, participatory decision-making

### 2022-05-02 Proposal/ch-experiment.tex
Move remainder of proposal to subfolder.

\section{Introduction}

@@

@@

The initial pilot study was conducted on December 21, 2018,
before the custom online platform was developed.
In this pilot, 18 participants engaged in a synchronous 60-minute deliberation.
Participants sat at individual computer workstations, separated by dividers.
Participants discussed the following question:
``Where should the use of electric scooters be permitted?''
Participants were placed in a single treatment group, and divided into 4
pods of size 4--5.
These pods were reassigned using the random-pod method for a total of three
rounds.
The deliberation took place in chat rooms running the free and open-source
OTree online experiment software \cite{chen_oTree_2016}.

@@

For the purposes of this dissertation, the preference profiles will be used
to quantify agreement and conflict over the course of deliberation.
Within social choice theory, agreement can be defined by a {\em consensus class},
a set of preference profiles meeting one of several consensus criteria
\cite{elkind_rationalizations_2016}.
Examples include strong unanimity (of rank orders),
unanimity (of winners),
majority (existence of majority),
Condorcet (existence of Condorcet winner),
and transitivity (of social preference).
On the most restrictive end, in strong unanimity and unanimity,
all decision-makers prefer the same alternative.
In contrast, transitivity only requires that the social preferences induced by
individual preferences are transitive.
Distance-rationalizable voting systems are defined by projecting the preference
profile onto the nearest consensus profile according to a suitable distance.
For example, Dodgson's method \cite{dodgeson_method_1876, brandt_computational_2012}
defines a distance based on swapping adjacent entries in individual preference profiles.
However, for most voting systems,
finding the consensus profile that minimizes distance is NP-Hard
\cite{elkind_rationalizations_2016}.
Social choice theory also provides measures of distances between two individual
preference rankings.
These measures can be extended to create measures of agreement for entire
preference profiles that can be calculated efficiently.

>>>>

\section{Introduction}
Networks of small interlocking pods appear as a common thread throughout
successful large-scale collaborations.
However, it remains unclear whether that network structure plays a causal role
in the success of those collaborations,
and whether the specific structural properties of that interlocking network matter.
Observational study suggests a correlation between properties such as degree,
structural efficiency, and structural inequality and the performance/productivity
of collaborations \cite{platt_network_2018}.
Numerical simulation suggests a causal relationship in a simplified theoretical
setting.
This chapter proposes an online experiment to bridge these two findings.

@@

In order to conduct studies of network deliberation online, I have modified the
free and open-source Loomio platform \cite{jackson_open_2016}
to implement network deliberation.
Loomio is a widely-used online deliberation platform which provides both
forum functionality (Figure \ref{fig:reply})
and various forms of voting, including ranked-choice (Figure \ref{fig:vote}).
Loomio is popular among grassroots organizations and worker-owned cooperatives
(Loomio itself is a worker-owned cooperative).
Loomio was chosen for its combination of built-in functionality,
existing user base, and open-source extensibility.

The modified platform implements both random-pod and long-path assignment methods
and automatically reassigns participants when the stage is advanced.
In addition to network deliberation features, I have made several modifications to implement
the experimental protocol.
Upon signing in for the first time,
participants are randomly assigned to a treatment group.
Participants are also assigned an alias which they will use throughout the
deliberation (Figure \ref{fig:alias}).

\begin{figure}
\label{fig:reply}
\center
\frame{\includegraphics[width=5in]{fig/Experiment/fig-reply.png}}
\caption{Participants can perform standard forum actions, such as posting and commenting.}
\end{figure}

\begin{figure}
\label{fig:vote}
\center
\frame{\includegraphics[width=3in]{fig/Experiment/fig-vote.png}}
\caption{Participants are presented with a ranked-choice voting screen between rounds of deliberation.}
\end{figure}

\begin{figure}
\label{fig:alias}
\center
\frame{\includegraphics[width=3in]{fig/Experiment/fig-alias.png}}
\caption{Participants are assigned an alias for the duration of the experiment.}
\end{figure}

@@

The initial pilot study was conducted on December 21, 2018,
before the custom online platform was developed.
In this pilot, 18 participants engaged in a synchronous 60-minute deliberation.
Participants sat at individual computer workstations, separated by dividers.
Participants discussed the following question:
``Where should the use of electric scooters be permitted?''
Participants were placed in a single treatment group, and divided into 4
pods of size 4--5.
These pods were reassigned using the random-pod method for a total of three
rounds.
The deliberation took place in chat rooms running the free and open-source
OTree online experiment software \cite{chen_otreeopen-source_2016}.

@@

For the purposes of this dissertation, the preference profiles will be used
to quantify agreement and conflict over the course of deliberation.
Within social choice theory, agreement can be defined by a {\em consensus class},
a set of preference profiles meeting one of several consensus criteria
\cite{elkind_rationalizations_2016}.
Examples include strong unanimity (of rank orders),
unanimity (of winners),
majority (existence of majority),
Condorcet (existence of Condorcet winner),
and transitivity (of social preference).
On the most restrictive end, in strong unanimity and unanimity,
all decision-makers prefer the same alternative.
In contrast, transitivity only requires that the social preferences induced by
individual preferences are transitive.
Distance-rationalizable voting systems are defined by projecting the preference
profile onto the nearest consensus profile according to a suitable distance.
For example, Dodgson's method \cite{dodgson_method_1876, brandt_computational_2012}
defines a distance based on swapping adjacent entries in individual preference profiles.
However, for most voting systems,
finding the consensus profile that minimizes distance is NP-Hard
\cite{elkind_rationalizations_2016}.
Social choice theory also provides measures of distances between two individual
preference rankings.
These measures can be extended to create measures of agreement for entire
preference profiles that can be calculated efficiently.
